\Appendix{Episode Description in the Emergency Response Scenario}

\section{Episode 1: Goal activation} % (fold)
\label{sec:episode_1_goal_activation}
\footnotesize
{
	\begin{longtable}{>{\raggedright}p{1.2in}>{\raggedright}p{1.2in}>{\raggedright}p{3in}}
\toprule 
\textbf{Actor} & \textbf{Type} & \textbf{Description}\tabularnewline
\midrule 
Fireman & Generate event (E1) & E1: A new victim has been found at a given location within the impacted
area\tabularnewline
\midrule 
\multirow{3}{1.2in}{Victim manager (VM)} & Consume event (E1) & Upon receiving E1, VM activate the goal that the victim needs to be
rescued, and decomposes the action to rescue the victim into sub-actions.\tabularnewline
\cmidrule{2-3} 
 & Perform action & In order to rescue the victim, VM first evaluates the status of the
victim. 

Because the radiation level of the victim exceeds the threshold, the
victim needs to be first decontaminated.

Because the victim is also physically injured, the victim needs to
be treated at a medical station after decontamination.

After the treatment, the victim needs to be evacuated to a shelter
for recovery.\tabularnewline
\cmidrule{2-3} 
 & Generate events (E2, E3, E4) & E2: A new goal to decontaminate the victim is activated

E3: A new goal to treat the victim is activated

E4: A new goal to evacuate the victim is activated\tabularnewline
\midrule 
\multirow{3}{1.2in}{Decontamination manager (DM)} & Consume event (E2) & Upon receiving E2, DM infers that the action to decontaminate is within
her responsibility and she is capable of performing it, so she is
committed to performing the action, decompose the action into sub-actions.\tabularnewline
\cmidrule{2-3} 
 & Perform action & To perform decontamination, DM first needs to assign a decontamination
station to the victim.

After the assignment of station (DS1) to the victim, the DM activates
the goal to transport the victim to the station. \tabularnewline
\cmidrule{2-3} 
 & Generate event (E5, E6) & E5: The station (DS1) has been assigned to the decontamination action
on victim.

E6: A new goal to transport the victim to the assigned decontamination
station is activated\tabularnewline
\midrule 
Medical manager (MM) & Consume event (E3) & Upon receiving E3, MM infers that the action to treat the victim is
within her responsibility. However, because the action depends on
the action to decontaminate the victim to be completed first, she
has to wait for now.\tabularnewline
\midrule 
\multirow{4}{1.2in}{Transportation manager (TM)} & Consume event (E4) & Upon receiving E4, TM infers that the action to evacuate the victim
is within her responsibility.

However, because the action depends on the action to treat the victim
to be completed first, she has to wait for that first.\tabularnewline
\cmidrule{2-3} 
 & Consume event (E6) & Upon receiving E5, TM infers that the action to transport the victim
to the station is within her responsibility and she is capable of
performing it, so she is committed to performing the action, and further
decompose the action.\tabularnewline
\cmidrule{2-3} 
 & Perform action & TM first needs to assign a driver (DR1) with a rescue vehicle to the
action.

Because the actual transportation action is out of the DM's responsibility,
she is waiting for the driver to perform it.\tabularnewline
\cmidrule{2-3} 
 & Generate event (E7) & E7: The driver (DR1) has been assigned to the action to transport
the victim \tabularnewline
\midrule 
Driver (DR1) & Consume event (E7) & Upon receiving E6, DR1 infers that the action to transport the victim
to the station is within her responsibility and she is capable of
performing it, so she starts to perform the action.\tabularnewline
\bottomrule
\caption{Episode 1 in the emergency response scenario}
\label{tab:episode_1_appendix}
\end{longtable}
}

% section episode_1_goal_activation (end)

\section{Episode 2: Plan development} % (fold)
\label{sec:episode_2_plan_development}
\footnotesize
{
	\begin{longtable}{>{\raggedright}p{1.2in}>{\raggedright}p{1.2in}>{\raggedright}p{3in}}
\toprule 
\textbf{Actor} & \textbf{Type} & \textbf{Description}\tabularnewline
\midrule 
Operator at the decontamination station & Generate event (E1) & E1: The quantity of available rinse tanks in the station (DS1) is
running low.\tabularnewline
\midrule 
\multirow{3}{1.2in}{Decontamination manager (DM)} & Consume event (E1) & Upon receiving E1, DM infers that the station (DS1) cannot work properly
due to this resource shortage. In order to ensure the victim is successfully
decontaminated, a new action to resolve the resource shortage is added
to the current plan to rescue the victim. \tabularnewline
\cmidrule{2-3} 
 & Perform action & DM first needs to locate an inventory where the rinse tanks can be
supplied.

After identification of the inventory (INV1), the DM activates the
goal to deliver the resource from the inventory to the station. \tabularnewline
\cmidrule{2-3} 
 & Generate event (E2) & E2: A new goal to deliver the resource from (INV1) to the station
(DS1) is activated.\tabularnewline
\midrule 
\multirow{3}{1.2in}{Inventory manager (IM)} & Consume event (E2) & Upon receiving E2, IM starts a new action to deliver the resource.\tabularnewline
\cmidrule{2-3} 
 & Perform action & IM first needs to prepare the resource for delivery. Once the resource
is ready, she activates the goal to transport the resource to the
station (DS1)\tabularnewline
\cmidrule{2-3} 
 & Generate event (E3) & E3: A new goal to transport the resource to the station (DS1) is activated\tabularnewline
\midrule 
\multirow{3}{1.2in}{Transportation manager (TM)} & Consume event (E3) & Upon receiving E3, TM starts a new action to transport the resource.\tabularnewline
\cmidrule{2-3} 
 & Perform action & TM first needs to assign a driver (DR2) with a rescue vehicle to the
action.

Because the actual transportation action is out of the DM's responsibility,
she is waiting for the driver to perform it.\tabularnewline
\cmidrule{2-3} 
 & Generate event (E4) & E4: The driver (DR2) has been assigned to the action to transport
the victim \tabularnewline
\midrule 
Driver (DR2) & Consume event (E4) & Upon receiving E4, DR2 infers that the action to transport the victim
to the station is within her responsibility and she is capable of
performing it, so she starts to perform the action.\tabularnewline
\bottomrule
\caption{Episode 2 in the emergency response scenario}
\label{tab:episode_2_appendix}
\end{longtable}
}
% section episode_2_plan_development (end)

\section{Episode 3: Responsibility transfer} % (fold)
\label{sec:episode_3_responsibility_transfer}
\footnotesize
{
\begin{longtable}{>{\raggedright}p{1.2in}>{\raggedright}p{1.2in}>{\raggedright}p{3in}}
\toprule 
\textbf{Actor} & \textbf{Type} & \textbf{Description}\tabularnewline
\midrule 
Driver (DR1) & Generate event (E1) & E1: The rescue vehicle with the driver (DR1) encounters a mechanical
breakdown.\tabularnewline
\midrule 
\multirow{3}{1.2in}{Transportation manager (TM)} & Consume event (E1) & Upon receiving E1, TM infers that because of the vehicle breakdown,
the DR1's action to transport the victim cannot be achieved.\tabularnewline
\cmidrule{2-3} 
 & Perform action & To solve the problem, TM attempts to assign another driver to the
action.

However, because all the drivers are currently in duty, the TM cannot
find a driver to perform the action until a later time.\tabularnewline
\cmidrule{2-3} 
 & Generate event (E2) & E2: The action to transport the victim has to be delayed because of
the unavailability of drivers.\tabularnewline
\midrule 
\multirow{2}{1.2in}{Decontamination manager (DM)} & Consume event (E2) & Upon receiving E2, DM infers that because of the delay to transport
the victim, the action to decontaminate the victim is also delayed.\tabularnewline
\cmidrule{2-3} 
 & Generate event (E3) & E3: The action to decontaminate the victim will be delayed because
of the delay of the transportation task.\tabularnewline
\midrule 
\multirow{2}{1.2in}{Victim manager (VM)} & Consume event (E3) & Upon receiving E3, VM infers that because of the delay to transport
the victim, the action to rescue the victim is delayed.\tabularnewline
\cmidrule{2-3} 
 & Generate event (E4) & E4: The action to rescue the victim is delayed because of the delay
of the transportation task.\tabularnewline
\midrule 
\multirow{2}{1.2in}{Fireman} & Consume event (E4) & Upon receiving E4, the fireman infers that although delivery of the
victim is not in the scope of her responsibility, however she can
help the rescue action by sending the victim to the decontamination
station (DS1). \tabularnewline
\cmidrule{2-3} 
 & Generate event (E5) & E5: The fireman activates the goal to send the victim to the decontamination
station (DS1). \tabularnewline
\midrule 
Transportation manager (TM) & Consume event (E5) & Upon receiving E5, TM infers that because of the new action of the
fireman, the responsibility of transporting the victim is now transferred
from her to the fireman.\tabularnewline
\bottomrule
\caption{Episode 3 in the emergency response scenario}
\label{tab:episode_3_appendix}
\end{longtable}
}
% section episode_3_responsibility_transfer (end)

\section{Episode 4: Opportunistic re-planning} % (fold)
\label{sec:episode_4_opportunistic_re_planning}
\footnotesize
{
	\begin{longtable}{>{\raggedright}p{1.2in}>{\raggedright}p{1.2in}>{\raggedright}p{3in}}
\toprule 
\textbf{Actor} & \textbf{Type} & \textbf{Description}\tabularnewline
\midrule 
Fireman & Generate event (E1) & E1: The physical injury of the victim becomes severe and prevents
the fireman from moving the victim into the vehicle.\tabularnewline
\midrule 
\multirow{2}{1.2in}{Victim manager (VM)} & Consume event (E1) & Upon receiving E1, VM evaluates the situation of the victim and decides
that because of the serious injury, the original plan to rescue the
victim is no longer appropriate. A new plan needs to be adopted: a
special team needs to be dispatched to the victim's location to decontaminate
and treat the victim on spot, and then the victim is directly sent
to the hospital for further treatment.\tabularnewline
\cmidrule{2-3} 
 & Generate events (E2, E3, E4, E5, E6) & E2: The current goal to decontaminate the victim is deactivated

E3: The current goal to treat the victim is deactivated

E4: The current goal to evacuate the victim is deactivated

E5: A new goal to dispatch a special team to the victim is activated

E6: A new goal to send the victim to the hospital is activated\tabularnewline
\midrule 
\multirow{4}{1.2in}{Decontamination manager (DM)} & Consume event (E2) & Upon receiving E2, DM removes the action to decontaminate the victim
from her current responsibility.\tabularnewline
\cmidrule{2-3} 
 & Consume event (E5) & Upon receiving E5, DM infers that in order to perform the action to
dispatch a special team to the victim, a decontamination operator
needs to be assigned to the team sent to the victim. \tabularnewline
\cmidrule{2-3} 
 & Perform action & By reviewing the available operators, DM decides on the operator that
will be sent to the victim\tabularnewline
\cmidrule{2-3} 
 & Generate event (E7) & The decontamination operator has been assigned to the special team
to the victim.\tabularnewline
\midrule 
\multirow{4}{1.2in}{Medical manager (MM)} & Consume event (E3) & Upon receiving E3, MM removes the action to treat the victim from
her current responsibility.\tabularnewline
\cmidrule{2-3} 
 & Consume event (E5) & Upon receiving E5, DM infers that in order to perform the action to
dispatch a special team to the victim, a paramedic needs to be assigned
to the team sent to the victim. \tabularnewline
\cmidrule{2-3} 
 & Perform action & By reviewing the available actors, DM decides on the paramedic that
will be sent to the victim\tabularnewline
\cmidrule{2-3} 
 & Generate event (E8) & The paramedic has been assigned to the special team to the victim.\tabularnewline
\midrule 
\multirow{5}{1.2in}{Transportation Manager} & Consume event (E4) & Upon receiving E4, TM removes the action to evacuate the victim from
her current responsibility.\tabularnewline
\cmidrule{2-3} 
 & Consume event (E5) & Upon receiving E5, TM infers that in order to perform the action to
dispatch a special team to the victim,  the action to transport the
team to the victim needs to be activated. \tabularnewline
\cmidrule{2-3} 
 & Perform action & TM first needs to assign a driver with a rescue vehicle to the action.
Because the deactivation of the original decontamination action, the
driver (DR2) who was assigned for delivering the equipment is now
available for this action.\tabularnewline
\cmidrule{2-3} 
 & Generate event (E9) & E9: The driver (DR2) has been assigned to the action to transport
the special team to the victim \tabularnewline
\cmidrule{2-3} 
 & Consume event (E6) & Upon receiving E6, TM infers that the action to send the victim to
hospital is within her responsibility.

However, because the action depends on other actions to be completed
first, she has to wait for those first.\tabularnewline
\midrule 
Driver (DR2) & Consume event (E9) & Upon receiving E9, DR2 starts to perform the action to transport the
special team to the victim.\tabularnewline
\bottomrule
\bottomrule
	\caption{Episode 4 in the emergency response scenario}
	\label{tab:episode_4_appendix}
\end{longtable}
}

% section episode_4_opportunistic_re_planning (end)