% Place acknowledgments below.
I would like to thank my parents for their support and encouragement, across the thousands of miles. Sharing every achievement with them along the way has made the journey of my Ph.D. immensely enjoyable.

My deepest gratitude to my advisor, Dr. Guoray Cai, for the time and effort he invested in me. Without him, I would have never come this far in my PhD research. And I really appreciate his encouragement and patience throughout the years, not to mention his excellent guidance and mentorship.

Special thanks to Dr. Alan M. MacEachren for being the external member of my thesis committee, for providing numerous suggestions and feedbacks to my dissertation, and for financially supporting my graduate research. I would also like to thank Dr. Mary Beth Rosson and Dr. Xiaolong (Luke) Zhang for serving on my thesis committee and for their insightful comments about my work. 

I learned a lot from my colleagues at the College of IST at Penn State. My extended network of graduate colleagues and friends were a great resource in conducting and writing up my research. 

Finally, I would like to express my gratitude to my beloved family. My wife, Lijun Xu, has been there every step of the way. I cannot thank enough for her patience, support, and belief in me and my work. I am very lucky to have her in my life. I would like to thank my son, Dylan Yu, for bringing me good luck and happiness. Hope I have done something that he would be proud of and encourages him to achieve things bigger.