%!TEX root = ../BoYu-Dissertation.tex
\subsubsection*{Education} % (fold)
\label{ssub:education}
\begin{itemize}
	\item PhD, Information Sciences and Technology at Pennsylvania State University, PA, USA, Aug. 2006 - May. 2013
	\item MS, Cartography and Geographical Information Systems at Peking University, China, Sep. 2003 - June 2006
	\item BS, Geographical Information Systems at Peking University, China, Sep. 1999 - June 2003
\end{itemize}
% subsubsection education (end)

\subsubsection*{Experience} % (fold)
\label{ssub:experience}
\begin{itemize}
	\item Collaboration Researcher, GE Global Research, USA, Dec. 2012 - Present
	\item Graduate Research/Teaching Assistant, College of IST, Penn State University, USA, Aug. 2006 - Dec. 2012
	\item Software Engineering Intern at Baidu.COM, Inc., Beijing, China, Sep. 2005 - Apr. 2006
	\item Graduate Research Assistant, Institute of Remote Sensing and GIS, Peking University, China, Sep. 2003 - June 2006  
\end{itemize}
% subsubsection experience (end)

\subsubsection*{Selected Publication} % (fold)
\label{ssub:selected_publication}
\begin{itemize}
	\item Guoray Cai, Bo Yu, and Dong Chen (2013). Modeling and Communicating the Conceptual Intent of Geo-analytical Tasks for Human-GIS Interaction. \emph{Transactions in GIS, S1(17), 2013}.
	\item Bo Yu and Guoray Cai (2012). Coordination of Emergency Response Operations via Event-Based Awareness Mechanism. \emph{2012 Information Systems for Crisis Response and Management (ISCRAM) conference, Vancouver, Canada}.
	\item Bo Yu and Guoray Cai (2010). Using Intentions and Plans of Mobile Activities to Guide Geospatial Web Service Composition. \emph{2010 IEEE Asia-Pacific Services Computing Conference, Hangzhou, China}.
	\item Guoray Cai and Bo Yu (2009). Spatial Annotation Technology for Public Deliberation. \emph{Transactions in GIS, S1(13), 2009, 123-146}.
	\item Bo Yu and Guoray Cai (2007). A Query-Aware Document Ranking Method for Geographic Information Retrieval. \emph{The fourth workshop on Geographic Information Retrieval, CIKM 07, Lisbon, Portugal. 2007}.
\end{itemize}
% subsubsection selected_publication (end)



