%!TEX root = ../BoYu-Dissertation.tex
\graphicspath{{Figures/}}

\chapter{Our Approach: Mediating Individual Awareness Processes} % (fold)
\label{cha:mediate_individual_awareness_processes}

supporting perception

support comprehension

support projection

\section{The Basic Interaction Scheme: Publish/Subscribe} % (fold)
\label{sec:the_basic_interaction_scheme_publish_subscribe}

% section the_basic_interaction_scheme_publish_subscribe (end)

\section{Existing Notification Approaches} % (fold)
\label{sec:existing_notification_approaches}
\subsection{Topic-Based Approach} % (fold)
\label{sub:topic_based_approach}

% subsection topic_based_approach (end)

\subsection{Type-Based Approach} % (fold)
\label{sub:type_based_approach}

% subsection type_based_approach (end)

\subsection{Content-Based Approach} % (fold)
\label{sub:content_based_approach}

% subsection content_based_approach (end)

% section existing_notification_approaches (end)

\section{Activity-Based Approach} % (fold)
\label{sec:activity_based_approach}
\subsection{Types of Events} % (fold)
\label{sub:types_of_events}
\subsubsection{External and internal events} % (fold)
\label{ssub:external_and_internal_events}
To understand events in geo-collaboration, an important distinction has to be made between \emph{external} events (in the environment) and internal events (in the activities). 

External events are changes in the physical environment that have impact on the performance of collaborative activities. For instance, the occurrence of a traffic accident may block the traffic flow, which makes an actor’s activity of delivering equipment to a medical station impossible to finish on time. External events in the environment can be characterized in three major categories: changes of identity (e.g. the occurrence of a traffic accident), attribute changes of an object (e.g. the contamination level of the chemical plan is increased), or spatial changes (e.g. the impacted area is enlarged due to the wind condition).

Internal events are state changes on the basic elements (i.e. resources, goals, activities) of geo-collaboration. Resource events indicate changes on the state of any resources that are used in the activities. Resource events can impact response activities in different ways: (1) the creation of a new resource can trigger new activities (for example, a new victim may lead to the activity to perform decontamination and medical treatment); (2) the state change of a resource can impact the activity that requires the use of the resource (e.g. the mechanical breakdown of a vehicle makes the delivery activity unable to complete); (3) the state change of a resource can enable/disable the activity that depends on it (e.g. a victim’s location change leads to the satisfaction of a precondition that the victim must be located at the medical station, which further enables the activity to perform medical treatment on the victim). Condition events indicate changes on a certain relationship between multiple objects in a collaborative activity. For example, a condition may express a spatial relationship that the victim must be located at the decontamination station. The condition will change the state from open to holding when a driver has picked up the victim and transpored him/her to the station. Activity events are described as change of activity states, such as the initiation of a new activity, the completion or delay of an ongoing activity. 

The distinction between external and internal events is very important to identify the set of events that should be captured and modeled in the awareness mechanism. Not all the external events in the environment are important for the actors to perform their activities. Rather, only a subset of the external events that can lead to changes within the activities (i.e. the internal events) is meaningful. The derivation of internal events from external events is an active cognitive process that requires interpreting the meaning of external events within the context of the collaborative activities.
% subsubsection external_and_internal_events (end)

\subsubsection{Local and remote events} % (fold)
\label{ssub:local_and_remote_events}
The distinction between external and internal events is based on the boundary between the surrounding environment and the overall collaborative activities as a whole. However, in order to understand the awareness information needed for each individual, another important distinction between local and remote events has to be made. 

The local and remote events are defined with aspect to the local scope of work for each individual actor. Local events reflect the state changes of basic elements (i.e. resources, goals, activities) within an actor’s local scope of work. For instance, the report of a new victim that needs to be decontaminated, the exceeding capacity of a decontamination station are local events within the decontamination manager’s local scope of work. Remote events are the changes outside an actor’s local scope of work. For instance, a victim that doesn’t need to be decontaminated is transported to a shelter is a remote event for the decontamination manager. The distinction of local and remote events is relative to each individual actor. A local event of one actor may be a remote event for another actor due to their different local scopes of work. In addition, a remote event can be internal, reflecting the state changes of another actor’s activities, or it can be external, reflecting changes in the environment. 

The distinction between local and remote events is important to identify the relevance of events that should be notified to each actor. Local events reflect changes within an actor’s local scope of work and therefore are directly relevant to the actor. On the other hand, due to the existence of the web of dependencies among activities, some of the remote events can (or potentially can) lead to changes in the actor’s local scope of work, i.e. the impact of a remote event may be propagated to the local scope of the actor as derived local events, and therefore they also become relevant. For instance, the traffic jam on the road is a remote event for the decontamination manager. However, because the traffic jam happens on the way of a victim to be delivered to a decontamination station, it can potentially delay the decontamination operation on the victim, which becomes a relevant local event for the decontamination manager. As a result, the goal of an awareness system is to notify an actor not only all the relevant local events, but also the subset of remote events that can be propagated to the actor’s local scope through the web of dependencies. 

% subsubsection local_and_remote_events (end)
% subsection types_of_events (end)

\subsection{Event Subscription} % (fold)
\label{sub:event_subscription}
\subsubsection{Specifying Local Scopes} % (fold)
\label{ssub:specifying_local_scopes}

% subsubsection specifying_local_scopes (end)

\subsubsection{Defining Event Patterns} % (fold)
\label{ssub:defining_event_patterns}

% subsubsection defining_event_patterns (end)
% subsection event_subscription (end)

\subsection{Event Matching} % (fold)
\label{sub:event_matching}

% subsection event_matching (end)

% section activity_based_approach (end)

\section{A Simulation Experiment} % (fold)
\label{sec:a_simulation_experiment}
We conduct a simulation experiment to compare the content-based event notification approach and the proposed activity-based approach. The basic hypothesis is that, in dynamic environment where the user's activity changes frequently, the proposed activity-based approach provides a more accurate way for the user to express their awareness interests than the content-based approach. We use the emergency response scenario as described in the Introduction Chapter to perform the experiment, from the perspective of the decontamination manager.

\subsection{Variables of Interests} % (fold)
\label{sub:variables_of_interests}
We are interested in the capability of these two notification approaches in handling different types of activity dynamics. Particularly, we are interested in three types of dynamics that may occur in the scenario: 
\begin{enumerate}
	\item \textbf{Parameter Change}: the value of a parameter is changed. In the scenario, it can be the case when the decontamination manager re-assign a new station where a victim will be decontaminated.
	\item \textbf{Plan Development}: the plan of the actor is changed. In the scenario, it can be the case when the decontamination manager starts to concern about certain equipment delivery after the stock is running low.
	\item \textbf{Local Scope Change}: the actor's local scope of work is changed. In the scenario, it can be the case when another manager joins the activity, and take a subset of tasks away from the modeling manager.  
\end{enumerate}

To perform the experiment, we define four scenes to reflect the three types of activity dynamics:
\begin{enumerate}
	\item Initial scene ($S_0$): indicates the current state of the decontamination manager's activities
	\item Parameter change scene ($S_1$): indicates one of the parameters of the decontamination manager (station) has been assigned with a new value, comparing with $S_0$
	\item Plan development scene ($S_2$): indicates one of the conditions has been elaborated into a more concrete plan, comparing with $S_1$
	\item local scope change ($S_3$): indicates the another actor joins the activity, and the local scope of the modeled user has been reduced to a smaller set, comparing with $S_2$
\end{enumerate}
% subsection variables_of_interests (end)

\subsection{Simulating Event Generation} % (fold)
\label{sub:event_generation}
A set of events are randomly generated based on the whole group activities. Given the current PlanGraph model, the events can happen on any of the activity, resource, or condition node, and indicate any types of possible changes that are defined in Chapter 5.
% subsection event_generation (end)

\subsection{Creating Subscriptions} % (fold)
\label{sub:content_based_subscriptions}
Two sets of content-based subscriptions will be generated. 
\begin{enumerate}
	\item The first set of subscriptions are defined merely based on the current interest of the manager in the initial scene ($S_0$). This set only includes the events that are relevant to the manager's current activities. Therefore, it can be considered as the minimal subscription set of events, using the content-based notification approach ($CON_{min}$).
	\item The second set of subscriptions are defined based on considering all the three possible activities dynamics defined in scenes ($S_1$, $S_2$, $S_3$). It includes all the possible events that might be relevant to the manager in all the scenes. Because we attempt to include the maximum set of events that could be relevant for the manager across all the scenes, we define this content-based subscription as ($CON_{max}$).  
\end{enumerate}

In addition, the activity-based subscription will also be generated, following the steps described in previous section. The subscription includes the specification of local scope of the decontamination manager, his/her intentions and detailed event patterns on each node in the local scope. We define this activity-based subscription as $ACT$.
% subsection creating_subscriptions (end)

\subsection{Procedures} % (fold)
\label{sub:procedures}
The experiment includes four runs. 
\begin{enumerate}
	\item The first run $R_0$ is the human judgment on the relevance of the simulated event set to form the baseline result. The human analysts run through the list of simulated event set to decide on whether each event is relevant to the modeled decontamination manager, under the four scenes respectively. To reduce the subjective errors of human judgment, two analysts will perform the judgment separately and then discuss on the differences to form the final sets of events.
	\item In the second run $R_1$, the simulated event set will be fed into the matching program following  content-based notification approach, using the set of subscriptions defined as $CON_{min}$.
	\item In the third run $R_2$, the simulated event set will be fed into the matching program following  content-based notification approach, using the set of subscriptions defined as $CON_{max}$.
	\item In the fourth run $R_3$, the the simuated event set will be fed into the matching program following activity-based notification approach, using the set of subscriptions defined as $ACT$.
\end{enumerate}
% subsection procedures (end)

\subsection{Measures} % (fold)
\label{sub:measures}
Two measures will be used to compare the results of different notification approaches:
\begin{enumerate}
	\item \textbf{Miss rate} is defined as the ratio between number of events that are considered as relevant in $R_0$ but missing in the result of $R_i$ ($i \in \{1, 2, 3\}$), and the total number of events considered as relevant in $R_0$.
	\item \textbf{False alarm rate} is defined as the ratio between number of events that are considered as relevant in $R_i$ ($i \in \{1, 2, 3\}$), but missing in the result of $R_0$, and the total number of events considered as relevant in $R_i$ ($i \in \{1, 2, 3\}$).
\end{enumerate}

% subsection measures (end)

\subsection{Results} % (fold)
\label{sub:results}
Some results we expect from the experiment can be:

\begin{enumerate}
	\item Comparing with the context-based notification using subscriptions defined as $CON_{min}$, the activity-based notification has lower miss rate in the scenes $S_1$, $S_2$, $S_3$. The difference between miss rates of the two methods increases as the scene evolves.
	\item Comparing with the context-based notification using subscriptions defined as $CON_{max}$, the activity-based notification has lower false alarm rate in the scenes $S_1$, $S_2$, $S_3$. The difference between false alarm rates of the two methods decreases as the scene evolves.
\end{enumerate}

% subsection results (end)

\subsection{Discussion} % (fold)
\label{sub:discussion}

% subsection discussion (end)


% section a_simulation_experiment (end)
\section{The Cognitive Basic} % (fold)
\label{sec:the_cognitive_basic}
The cognitive principles that guide the design of visual displays for event interpretation.

\begin{enumerate}
	\item The schemata theory (Bartlett et al.): when human respond to incoming stimuli, they make use of their existing knowledge as a frame of reference to make sense of the incoming stimuli and produce behavior. The active organization of existing knowledge is defined as `schema'. Schemata can be merely mental templates in human mind, or retrieved dynamically from the external visual artifacts. 
	\item The relevance principle (Kosslyn 2006): Visual displays should present no more or no less information than is needed by the user. Presenting all of the relevant information in the display relieves the user of the need to maintain a detailed representation of this information in working memory, where presenting too much information in the display leads to visual clutter or distraction by irrelevant information (Rosenholtz et al. 2007; Wickens \& Carswell, 1995).
	\item The task specificity principle (Hegarty et al. 2009). Visual displays are used for many different tasks, and there is no such thing as a `best' visual display, independent of the task to be carried out with this display. 
\end{enumerate}

What these cognitive principles can inform us are:
\begin{enumerate}
	\item When the user interprets an awareness event, he/she needs to activate and connect the event to his/her contextual knowledge of the situation. The visual display can enhance the interpretation by externalizing part of the knowledge in the visual representation and freeing up working memory resources.
	\item However, the effectiveness of the visual display depends on how much relevant information is represented. We want to provide the users with just enough contextual information without causing visual clutter or distraction.
	\item The relevance of contextual information to be displayed is dependent on the current task to be carried out, or the decision that need to be made by the user.

\end{enumerate}
% section the_cognitive_basic (end)

\section{Activity-Aware Event Interpretation} % (fold)
\label{sec:activity_aware_event_interpretation}
The problem we want to address here is that: how the computational model of activities and local scopes can inform the system to decide on what contextual information should be displayed when the users interpret awareness events.

The decisions can be made based on two factors: the incoming event and the activities/local scope the user is working on. By maintaining the PlanGraph model, the system can infer how the event will impact the user's activities, which can be used to infer what decisions the user needs to make next, or what tasks the user will focus on, which can be used to decide on what should be displayed on the user interface.

Generate the set of rules based on SharedPlan Theory.

Some examples:

1. If the event initiates a new goal of the user, by elaborating on the plan, the system can infer that the user will work on identifying the set of parameters first.

2. If the event indicates the failure of a parameter, and there is no plan to fix the failure, the system can infer that the user will re-assign the values of this parameter.

3. if the event indicates the failure of a parameter, and there is plan to fix the failure, the system can infer that the user will work on the plan to fix it.

% section activity_aware_event_interpretation (end)

\section{Experimental Study}
\subsection{Hypotheses} % (fold)
\label{sec:hypotheses}
The general assumption is that maps with contextual information that is adapted to the user's current activities are more effective to support awareness event interpretation, than maps with fixed contextual information.
% section hypotheses (end)

\subsection{Experimental Design} % (fold)
\label{sec:experimental_design}
\subsubsection{Participants} % (fold)
\label{sub:participants}
Twelve undergraduate or graduate students will be recruited as participants in this study. All participants need to use computers on a daily basis. Prior experience with emergency response planning or operations is welcomed, but not required.

% subsection participants (end)
\subsubsection{Tasks} % (fold)
\label{sub:tasks}
The participants are asked to perform simulation tasks in an emergency response scenario, from the perspective of the decontamination manager. The tasks that the participants need to perform are driven by the events they receive.

Some example events and triggered tasks include:

\begin{enumerate}
	\item E1: A new victim that needs to be decontaminated is reported by the victim manager. The task that the participant needs to do is to assign a decon station for the victim.
	\item E2: The impacted area is enlarged and a decon station becomes inside of the impacted area. The task that the participants need to do is to re-assign the victims that are assigned to the station to other stations.
	\item E3: A type of resource in a particular decon station is running low in stock. The participant need to request for a resource delivery from one of the suppliers.
	\item E4: The delay on a victim's arrival at assigned station. The participant needs to check whether it will conflict with the victim's deadline on decontamination. If so, he/she needs to plan for sending the victim to other station.
\end{enumerate}

% subsection task (end)

\subsubsection{Design settings} % (fold)
\label{sub:design_settings}
Participants are randomly assigned to one of the two design settings in one experiment session to perform their tasks. The three design settings are described as follow: 
\begin{itemize}
    \item The first design setting ($D_1$) shows a map showing locations of all the stations, victims, and resource suppliers as separate layers. The user can turn on/off each layer, or filter the objects based on spatial distances.
    
    \item In the second design setting ($D_2$), the features displayed in the map are adapted based on the event and the tasks that the user needs to perform. 
\end{itemize}
% subsection design_settings (end)

\subsubsection{Procedure} % (fold)
\label{sub:procedure}
In the beginning of each experiment session, the participant is randomly assigned to one of the three design settings and given a couple of minutes to learn the functions of the assigned design setting. In addition to the training of the simulation system, the participant is also provided with the general picture of the collaborative activity. Information about other agents, related resources in the activity, and the current goals and tasks of the group are presented to the participant, so he/she has a clear understanding of the task in the context of the whole activity and provides the basis for interpreting the awareness events. After the training session, the participant is asked to perform the simulation task. During the performance of the task, the participant is notified with a list of awareness events, and asked to finish the assessment form to answer relevant questions after each event notification.

Each experiment session is composed as a series of interaction episodes between the participant and the system. Each interaction episode defines how an agent (role-played by the participant) responds to a particular event in the collaborative scenario. Each interaction episode starts with the provision of an awareness event in the interface, and then the participant is perform corresponding tasks based on their understanding of the event.

% subsection procedure (end)

\subsubsection{Measurement} % (fold)
\label{sub:data_collection}
To measure the overall performance during the experiment session, we record the completion time of each experiment session and the participant's responses to each awareness event. In the design phase of the awareness events, each event is associated with a set of optimal responses. The participant's responses are then compared with the optimal set to indicate the quality level of task completion. 

% section data_collection (end)

\subsection{Results} % (fold)
\label{sec:results_and_discussion}
In general, we expect that 

1. the participants in design setting $D_2$ will generate more valid responses than the participants in design setting $D_1$.

2. the participants in design setting $D_2$ should spend less time to finish the tasks than the participants in design settings $D_1$ in average.
% section results_and_discussion (end)

\subsection{Discussion}



% chapter mediate_individual_awareness_processes (end)




 

